Local wind regime induced by the presence of giant linear dunes.

### 2022-01-03 Paper/main.tex
Reviewing all text + adding abstract and acknowledgments

\author{Cyril Gadal \and Pauline Delorme \and Clément Narteau \and
        Giles Wiggs \and Matthew Baddock \and Joanna M. Nield \and Philippe Claudin }

@@

\begin{abstract}

@@

  On the other hand, it depends on rigidity of the capping surface, as its deformation releases the confinement effect inducing streamline compression. This is typically quantified using the Froude number~\citep{Vosper2004, Stull2006, Sheridan2006, Hunt2006, Jiang2014}:
  \begin{equation}
        \mathcal{F} = \displaystyle\frac{U}{\sqrt{\displaystyle\frac{\Delta\rho}{\rho}gH}},
  \end{equation}
  where $U$ is the wind velocity at the top of the ABL, $\rho$ its average density, $\Delta\rho$ the density jump between the ABL and the FA.

@@

  The presence of a stratification in the free atmosphere can also impact the flow confinement, depending on its ability to deform under the presence of an underlying obstacle. This can be quantified using the internal Froude number~\citep{Vosper2004, Stull2006, Sheridan2006, Hunt2006, Jiang2014}:
  \begin{equation}
    \mathcal{F}_{\textup{I}} = \displaystyle\frac{k U}{N},
  \end{equation}
  where $N = \sqrt{-(g/\rho)(\partial \rho/\partial z)}$ is characteristic of the stratification. Both Froude numbers have the same qualitative effect on flow confinement, as they quantify the rigidity of the overlying layers. This is confirmed by figure~\ref{Fig12_supp}, where we can also find the different regimes related to wind disturbances described previously for the Froude number $\mathcal{F}$.

@@

\begin{acknowledgements}
[citing all grants ...]\\
Finally, we would like to acknowledge the contributors of the following open-source python librairies, Matplotlib \citep{Hunter2007}, Numpy \citep{Harris2020} and Scipy \citep{Virtanen2020}, which provide an incredibly efficient ecosystem allowing scientific research in Python.
\end{acknowledgements}

@@

The two components of the basal shear stress $\boldsymbol{\tau} = \rho u_{*}\boldsymbol{u}_{*}$, constant in a flat bottom situation, can then be written without loss of generality as:
\begin{align}
  \tau_{x} & = \tau_{0}\left(1 + k\xi_{0}\sqrt{\mathcal{A}_{x}^{2} + \mathcal{B}_{x}^{2}}\cos\left[k\left(\cos(\alpha)x + \sin(\alpha)y\right) + \phi_{x}\right]\right), \\
  \tau_{y} & = \tau_{0}k\xi_{0}\sqrt{\mathcal{A}_{y}^{2} + \mathcal{B}_{y}^{2}}\cos\left[k\left(\cos(\alpha)x + \sin(\alpha)y\right) + \phi_{y}\right],
\end{align}
where $\tau_{0}$ is the basal shear stress on a flat bed, and $\phi_{x, y} = \tan^{-1}\left(\mathcal{B}_{x, y}/\mathcal{A}_{x, y}\right)$. The in--phase and in--quadrature hydrodynamical coefficients $\mathcal{A}_{x, y}$ and $\mathcal{B}_{x, y}$ are functions of the flow conditions, i.e the bottom roughness, the vertical flow structure or the incident flow direction \citep{Fourriere2010, andreotti2009, Andreotti2012}.

@@

\paragraph{Sediment transport}
The prediction of sand fluxes from wind data has been a long standing issue in geomorphological studies~\citep{Fryberger79, Pearce2005, Sherman2012, Shen2019}. Based on laboratory studies in wind tunnels \citep{Rasmussen96, Iversen99, Creyssels2009, Ho2011}, as well as physical considerations \citep{Ungar1987, Andreotti2004bis, Duran2011, Pahtz2020}, it has been shown that the steady saturated sand flux over a flat sand bed depends linearly on the shear stress:
\begin{equation}
  \label{transport_law_quadratic}
  \frac{q_{\textup{sat, t}}}{Q} = \Omega\sqrt{\Theta_{\textup{th}}}\left(\Theta_{t} - \Theta_{\textup{th}}\right),
\end{equation}
where $\Omega$ is a proportionality constant, $Q = d\sqrt{(\rho_{\textup{s}} - \rho)gd/\rho}$ is a characteristic flux, $\Theta = \rho u_{*, t}^{2}/(\rho_{\textup{s}} - \rho)gd$ the Shields number, and $\Theta_{\textup{th}}$ its threshold value for incipient sediment transport. Here, $\rho_{\textup{s}} = 2.6~\textup{g}~\textup{cm}^{-3}$ and $d=180~\mu\textup{m}$ are the grain density and diameter, and $g$ is the gravitational acceleration.

@@

For each station, the hydrodynamic roughness is calibrated by finding the one that minimizes the relative difference $\delta$ between the wind vectors of both datasets:
\begin{equation}
  \label{metric_roughness}
      \delta = \frac{\sqrt{\langle\| \boldsymbol{u}_{*, \textrm{era}} - \boldsymbol{u}_{*, \textrm{station}} \|^{2}\rangle_{t}}}{\sqrt{ \langle \| \boldsymbol{u}_{*, \textrm{era}} \| \rangle_{t}\langle \| \boldsymbol{u}_{*, \textrm{station}} \| \rangle_{t}}}
\end{equation}

This $\delta$--parameter is computed for hydrodynamic roughness values ranging from $10^{-5}$~m to $10^{-2}$~m for the different stations. A shown by figure~\ref{Fig3_supp}, the minimum of $\delta$ in the space ($z_{0,\, \textup{Era}}$, $z_{0,\, \textup{in situ}}$) forms a line. We thus take the roughness of the Era5Land dataset as the typical value when sediment transport occurs, $10^{-3}$~m, corresponding to the thickness of the transport layer \citep{Duran2011}. It leads for the Adamax, Deep Sea, Huab and South Namib stations values of $2.7~\textup{mm}$, $0.76~\textup{mm}$, $0.12~\textup{mm}$ and $0.48~\textup{mm}$, respectively.

The choice of the hydrodynamic roughness values only impacts the calculated shear velocities, but note the wind directions. As such, most of our conclusions are then independent of such a choice, and only the magnitude of the wind velocity attenuation in confined situation might be affected.

\begin{figure}
  \centering
  \includegraphics[scale=1]{Figures/Figure3_supp.pdf}
  \caption{Calibration of the hydrodynamic roughnesses. The metric $\delta$ defined in \eqref{metric_roughness} is represented in colorscale as a function of the hydrodynamic roughnesses chosen for the Era5-Land and in situ datasets, for the Adamax (\textbf{a}), Deep Sea (\textbf{b}), Huab (\textbf{c}) and South Namib (\textbf{d}) Stations. The red dashed and plain lines shows the minima of $\delta$ and the identity line. The orange lines and dots highlight the chosen hydrodynamic roughnesses for the local datasets, by imposing $z_{0, \textup{\textup{ERA}}} = 1~\textup{mm}$, leading for each station to $2.7~\textup{mm}$, $0.76~\textup{mm}$, $0.12~\textup{mm}$ and $0.48~\textup{mm}$, respectively.}
  \label{Fig3_supp}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[scale=1]{Figures/Figure6_supp.pdf}
  \caption{Distributions of wind direction at the South Namib Station for the Era5Land climate reanalysis (orange) and the in situ measurements (blue). In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) in the Era5Land dataset. The gray vertical dashed lines indicate the dune orientation. The numbers at the top center give the percentage of time steps selected in each sub-range, as well as the percentage of them corresponding to the day (between 1000 UTC and 2200 UTC). The purple frame highlights the regime (small wind velocities, nocturnal summer wind) during which the wind data from both datasets differs. A similar figure can be obtained for the South Namib station (see Fig.~\ref{Fig3}).}
  \label{Fig6_supp}
\end{figure}

@@

In order to estimate the relevant non-dimensional numbers, one need to estimate in addition to the wind and dune properties some parameters of the ABL. The Era5 dataset provides a direct bulk estimate of the ABL depth $H$ from a bulk Richardson number calculation, as well as vertical profiles of the geopotential $\phi$, temperature $T$ and specific humidity $e_{\textup{w}}$ at given pressure levels $P$. From these quantities, the virtual potential temperature, which takes into account the vertical pressure and humidity changes, can be calculated as:
\begin{equation}
  T_{\textup{vp}} = T\left(1 + \left[R_{\textup{M}} - 1\right]e_{\textup{w}}\right)\left(\frac{P_{0}}{P}\right)^{P_{\textup{c}}(1 - 0.24e_{\textup{w}})},
\end{equation}
where $P_{0} = 10^{5}~\textup{Pa}$ is the standard pressure, $P_{\textup{c}} = 0.2854$ the Poisson coefficient for dry air and $R_{\textup{M}}=1.61$ is the ratio between the melocular masses of dry air and water. The vertical coordinates are calculated as:
\begin{equation}
  z = \frac{\phi R_{\textup{t}}}{g R_{\textup{t}} - \phi},
\end{equation}
where $R_{\textup{t}} = 6356766~\textup{m}$ is the average Earth radius, and $g = 9.81~\textup{m}~\textup{s}^{-2}$ the gravitational acceleration.

Example of obtained vertical profiles of the virtual potential temperature are shown in Fig.~\ref{Fig9_supp}. On each of them, an average is computed below the ABL depth given by the Era5 dataset, and a linear function is fitted above.

Under the Boussinesq approximation, the temperature variations are assumed to induce most of those of the density, leading to $\Delta\rho/\rho \simeq \Delta T_{\textup{vp}}/T_{\textup{vp}}$. Here, $T_{\textup{vp}}/T_{\textup{vp}}$ is the relative virtual potential temperature jump at the capping, directly measured on the vertical profiles.

Following \citet{Tritton2012}, the relative density jump at the capping layer

\begin{figure}
  \centering
  \includegraphics[scale=1]{Figures/Figure9_supp.pdf}
  \caption{\textbf{a}: Vertical profiles of the virtual potential temperature at 3 different time steps (blue - 29/11/2012 - 1100 UTC, orange - 21/03/2017 - 1200 UTC, green - 21/03/2017 - 2000 UTC) at the South Namib station. Dots: data from the ERA5 reanalysis. Transparent dashed lines: boundary layer height given by the ERA5 reanalysis, calculated from the bulk Richardson number \citep{seidel2012}. Plain lines: vertical (boundary layer) and linear (free atmosphere) fits to estimate the quantities in Fig.~\ref{Fig10_supp}. \textbf{b}: Examples of ill-processed vertical profiles at 3 different time steps (blue - 2/12/2013 - 2300 UTC, orange - 20/03/2017 - 0000 UTC, green - 14/07/2017 - 1400 UTC) at the South Namib station. \textbf{c}: Hourly distribution of ill-processed vertical profiles. \textbf{d}: Monthly distribution of ill-processed vertical profiles. \cg{These profiles are ill-processed because the temperature found at the boundary layer from the linear fit in the free-atm is smaller than the average one inside the boundary layer. This is an unstable situation, which does not allow to calculate the surface Froude number.}}
  \label{Fig9_supp}
\end{figure}

>>>>

\author{Cyril Gadal \and Pauline Delorme \and Clément Narteau \and
        Giles F.S. Wiggs \and Matthew Baddock \and Joanna M. Nield \and Philippe Claudin}

@@

\begin{abstract}
  In large-scale studies of arid areas, sediment fluxes and resulting geomorphological quantities (sediment pathways, erosion rates, dune morphodynamics) are often derived from wind data.
  %
  However, the feedback of nearby topography on the flow, such as kilometric giant dunes, is rarely taken into account because barely studied.
  %
  Here, we compare locally measured wind data to the predictions of the Era5Land climate reanalysis in four different places across and around the Namib sand sea. In places with flat environments, we show that the two wind datasets agree with each other. In contrast, within the sand sea, we show that they significantly differ due to the presence of giant linear dunes. We quantify this discrepancy, and link it to dynamics of the atmospheric boundary layer.
  %
  Aside setting limits to the applicability of the Era5land dataset, this study highlights the giant dunes feedback on the wind, and discuss its implications for smaller scale bedforms.
  %
  All codes used, from the raw data to the figures of the paper, are publicly available with full documentation at https://github.com/Cgadal/GiantDunes (will be made public upon acceptance of this manuscript for publication).

@@

  On the other hand, it depends on rigidity of the capping surface, as its deformation releases the confinement effect inducing streamline compression. This is typically quantified using the Froude number~\citep{Vosper2004, Stull2006, Sheridan2006, Hunt2006, Jiang2014}:
  \begin{equation}
        \mathcal{F} = \displaystyle\frac{U}{\sqrt{\displaystyle\frac{\Delta\rho}{\rho_{0}}gH}},
  \end{equation}
  where $U$ is the wind velocity at the top of the ABL, $\rho_{0}$ its average density, $\Delta\rho$ the density jump between the ABL and the FA.

@@

  The presence of a stratification in the free atmosphere can also impact the flow confinement, depending on its ability to deform under the presence of an underlying obstacle. This can be quantified using the internal Froude number~\citep{Vosper2004, Stull2006, Sheridan2006, Hunt2006, Jiang2014}:
  \begin{equation}
    \mathcal{F}_{\textup{I}} = \displaystyle\frac{k U}{N},
  \end{equation}
  where $N = \sqrt{-(g/\rho_{0})(\partial \rho/\partial z)}$ is characteristic of the stratification. Both Froude numbers have the same qualitative effect on flow confinement, as they quantify the rigidity of the overlying layers. This is confirmed by figure~\ref{Fig12_supp}, where we can also find the different regimes related to wind disturbances described previously for the Froude number $\mathcal{F}$.

@@

\begin{acknowledgements}
We would like to acknowledge the contributors of the following open-source python librairies, Matplotlib \citep{Hunter2007}, Numpy \citep{Harris2020} and Scipy \citep{Virtanen2020}, which provide an incredibly efficient ecosystem allowing scientific research in Python.

Era5 and Era5Land datasets are publicly available at the Copernicus Climate Change Service (C3S) Climate Data Store. The locally measured wind data can be found at \cg{upload on public data repository}. The digital elevation models from the Shuttle Radar Topography Mission are publicly available from Nasa servers, and can be downloaded at https://dwtkns.com/srtm30m/. Fully documented codes used to analyze this study are available at https://github.com/Cgadal/GiantDunes (will be made public upon acceptance of this manuscript for publication).

[citing all grants ...]\\
\end{acknowledgements}

@@

The two components of the basal shear stress $\boldsymbol{\tau} = \rho_{0} u_{*}\boldsymbol{u}_{*}$, constant in a flat bottom situation, can then be written without loss of generality as:
\begin{align}
  \tau_{x} & = \tau_{0}\left(1 + k\xi_{0}\sqrt{\mathcal{A}_{x}^{2} + \mathcal{B}_{x}^{2}}\cos\left[k\left(\cos(\alpha)x + \sin(\alpha)y\right) + \phi_{x}\right]\right), \\
  \tau_{y} & = \tau_{0}k\xi_{0}\sqrt{\mathcal{A}_{y}^{2} + \mathcal{B}_{y}^{2}}\cos\left[k\left(\cos(\alpha)x + \sin(\alpha)y\right) + \phi_{y}\right],
\end{align}
where $\tau_{0}$ is the basal shear stress on a flat bed, and $\phi_{x, y} = \tan^{-1}\left(\mathcal{B}_{x, y}/\mathcal{A}_{x, y}\right)$. The in--phase and in--quadrature hydrodynamical coefficients $\mathcal{A}_{x, y}$ and $\mathcal{B}_{x, y}$ are functions of the flow conditions, i.e the bottom roughness, the vertical flow structure or the incident flow direction \citep{Fourriere2010, andreotti2009, Andreotti2012}.

@@

\paragraph{Sediment transport}
The prediction of sand fluxes from wind data has been a long standing issue in geomorphological studies~\citep{Fryberger79, Pearce2005, Sherman2012, Shen2019}. Based on laboratory studies in wind tunnels \citep{Rasmussen96, Iversen99, Creyssels2009, Ho2011}, as well as physical considerations \citep{Ungar1987, Andreotti2004bis, Duran2011, Pahtz2020}, it has been shown that the steady saturated sand flux over a flat sand bed depends linearly on the shear stress:
\begin{equation}
  \label{transport_law_quadratic}
  \frac{q_{\textup{sat, t}}}{Q} = \Omega\sqrt{\Theta_{\textup{th}}}\left(\Theta_{t} - \Theta_{\textup{th}}\right),
\end{equation}
where $\Omega$ is a proportionality constant, $Q = d\sqrt{(\rho_{\textup{s}} - \rho_{0})gd/\rho_{0}}$ is a characteristic flux, $\Theta = \rho_{0} u_{*, t}^{2}/(\rho_{\textup{s}} - \rho_{0})gd$ the Shields number, and $\Theta_{\textup{th}}$ its threshold value for incipient sediment transport. Here, $\rho_{\textup{s}} = 2.6~\textup{g}~\textup{cm}^{-3}$ and $d=180~\mu\textup{m}$ are the grain density and diameter, and $g$ is the gravitational acceleration.

@@

As the regionally predicted and locally measured velocities are available at different heights, we can not compare them directly. We then convert all velocities into shear velocities $u_{*}$, characteristic of the turbulent velocity profile \citep{Spalding1961, Stull1988}:
\begin{equation}
  \frac{u(z)}{u_{*}} = \frac{1}{\kappa}\ln\left(\frac{z}{z_{0}}\right),
\end{equation}
where $z$ is the vertical coordinate, $\kappa \approx 0.4$ the von Kármán constant and $z_{0}$ the hydrodynamic roughness. Several field measurements of hydrodynamic roughnesses are available. In the absence of sediment transport, it scales with the geometric roughness of the topography \citep{Pelletier2016}. When transport occurs, it then scales with the thickness of the transport layer ($\sim 1$~mm), depending on the wind velocity and grain properties \citep{Sherman2008, Zhang2016, Field2018}.

The different environments of our stations (vegetated, arid, sandy) makes it difficult to develop a precise model for the computation of the hydrodyamic roughness. Furthermore, we do not have precise enough topographical measurements that would allow to compute the geometric roughness. We choose to leave aside this complexity, and use a different approach. The selected hydrodynamic roughness is then the one that leads to the best possible matching between the regionally predicted and locally measured winds.

For each station, the hydrodynamic roughness is calibrated by finding the one that minimizes the relative difference $\delta$ between the wind vectors of both datasets:
\begin{equation}
  \label{metric_roughness}
      \delta = \frac{\sqrt{\langle\| \boldsymbol{u}_{*, \textrm{era}} - \boldsymbol{u}_{*, \textrm{station}} \|^{2}\rangle_{t}}}{\sqrt{ \langle \| \boldsymbol{u}_{*, \textrm{era}} \| \rangle_{t}\langle \| \boldsymbol{u}_{*, \textrm{station}} \| \rangle_{t}}}
\end{equation}
This $\delta$--parameter is computed for hydrodynamic roughness values ranging from $10^{-5}$~m to $10^{-2}$~m for the different stations. Note that for the Deep Sea and South Namib stations, where the giant dunes feedback presumably affect the wind, we take into account the non-deflected winds only in the calculation of $\delta$ (with a $15\circ$ tolerance).

A shown by figure~\ref{Fig3_supp}, the minimum of $\delta$ in the space ($z_{0}^{\textup{Era5Land}}$, $z_{0}^{\textup{local}}$) forms a line. We thus take the roughness of the Era5Land dataset as the typical value when sediment transport occurs, $10^{-3}$~m, corresponding to the thickness of the transport layer \citep{Duran2011}. It leads for the Adamax, Deep Sea, Huab and South Namib stations values of $2.7~\textup{mm}$, $0.76~\textup{mm}$, $0.12~\textup{mm}$ and $0.48~\textup{mm}$, respectively.

The choice of the hydrodynamic roughness values impacts the calculated shear velocities only, but not the wind directions. As such, most of our conclusions are independent of such a choice. However, it may affect the magnitude of the wind velocity attenuation/amplification in flow confinement situations.

\begin{figure}
  \centering
  \includegraphics[scale=1]{Figures/Figure3_supp.pdf}
  \caption{Calibration of the hydrodynamic roughnesses. The metric $\delta$ defined in \eqref{metric_roughness} is represented in colorscale as a function of the hydrodynamic roughnesses chosen for the Era5-Land and local winds, for the (\textbf{a}) Adamax, (\textbf{b}) Deep Sea, (\textbf{c}) Huab and (\textbf{d}) South Namib stations. The red dashed and plain lines shows the minima of $\delta$ and the identity line, respectively. The orange lines and dots highlight the chosen hydrodynamic roughnesses for the local winds, by imposing $z_{0}^{\textup{Era5Land}} = 1~\textup{mm}$. It leads for for each station to $2.7~\textup{mm}$, $0.76~\textup{mm}$, $0.12~\textup{mm}$ and $0.48~\textup{mm}$, respectively.}
  \label{Fig3_supp}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[scale=1]{Figures/Figure6_supp.pdf}
  \caption{Distributions of wind direction at the South Namib Station for the Era5Land climate reanalysis (orange) and the local measurements (blue). In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) in the Era5Land dataset. The gray vertical dashed lines indicate the dune orientation. The numbers at the top center give the percentage of time steps selected in each sub-range, as well as the percentage of them corresponding to the day (between 1000 UTC and 2200 UTC). The purple frame highlights the regime (small wind velocities, nocturnal summer wind) during which the wind data from both datasets differs. A similar figure can be obtained for the South Namib station (see Fig.~\ref{Fig3}).}
  \label{Fig6_supp}
\end{figure}

@@

The estimation of the non-dimensional numbers requires the computation of meteorological quantities representative of the current atmospheric boundary layer. In arid areas, the vertical structure of the atmosphere can be approximated by well mixed convective boundary layer of height $H$, topped by the stratified free atmosphere \citep{Stull1988, Shao2008}.
%
In this type of structure, the virtual potential temperature is constant inside the boundary layer, and then increases linearly in the stratified free atmosphere (Fig.~\ref{Fig9_supp}a):
\begin{equation}
  T_{\textup{vp}}(z) = \begin{cases}
                          T_{0} &\text{for $z < H$},\\
                          T_{0}\left(1 + \displaystyle\frac{\Delta T_{\textup{vp}}}{T_{0}} + \frac{N^{2}}{g}(z - H)\right) &\text{for $ z > H$},
                        \end{cases}
\end{equation}
where, $T_{0}$ is the surface virtual potential temperature, $\Delta T_{\textup{vp}}$ the discontinuity at the capping layer and $N$ the Brunt-Väisälä frequency, characteristic of the stratification.

The Era5 dataset provides a direct estimate of the ABL depth $H$ from a bulk Richardson number calculation, as well as vertical profiles of the geopotential $\phi$, temperature $T$ and specific humidity $e_{\textup{w}}$ at given pressure levels $P$. From these quantities, the virtual potential temperature can be calculated as:
\begin{equation}
  T_{\textup{vp}} = T\left(1 + \left[R_{\textup{M}} - 1\right]e_{\textup{w}}\right)\left(\frac{P_{0}}{P}\right)^{P_{\textup{c}}(1 - 0.24e_{\textup{w}})},
\end{equation}
where $P_{0} = 10^{5}~\textup{Pa}$ is the standard pressure, $P_{\textup{c}} = 0.2854$ the Poisson coefficient for dry air and $R_{\textup{M}}=1.61$ is the ratio between the molecular masses of dry air and water. The vertical coordinates are calculated as:
\begin{equation}
  z = \frac{\phi R_{\textup{t}}}{g R_{\textup{t}} - \phi},
\end{equation}
where $R_{\textup{t}} = 6356766~\textup{m}$ is the average Earth radius, and $g = 9.81~\textup{m}~\textup{s}^{-2}$ the gravitational acceleration.

Example of obtained vertical profiles of the virtual potential temperature are shown in Fig.~\ref{Fig9_supp}a. On each of them, an average is computed below the ABL depth directly given by the Era5 dataset, and a linear function is fitted above. Note that under the Boussinesq approximation, the temperature variations are assumed to induce most of those of the density, leading to $\Delta\rho/\rho_{0} \simeq \Delta T_{\textup{vp}}/T_{0}$ (see supplementary material of \citep{andreotti2009}). Note that we removed some profiles that displayed a vertical structure that could not be approximated by the simple model used here, leading for example to negative values of $\Delta T_{\textup{vp}}$ (see Fig.~\ref{Fig9_supp}b). While these profiles dominantly occurs in winter, when the two winds blows, they are evenly spread across the hours of the day, and represent 12~\% of our data only (see Fig.~\ref{Fig9_supp}c--d). For these reasons, we are confident this does not affect pour conclusions.

\begin{figure}
  \centering
  \includegraphics[scale=1]{Figures/Figure9_supp.pdf}
  \caption{\textbf{a}: Vertical profiles of the virtual potential temperature at 3 different time steps (blue - 29/11/2012 - 1100 UTC, orange - 21/03/2017 - 1200 UTC, green - 21/03/2017 - 2000 UTC) at the South Namib station. Dots: data from the ERA5 reanalysis. Transparent dashed lines: boundary layer height given by the ERA5 reanalysis, calculated from the bulk Richardson number \citep{seidel2012}. Plain lines: vertical (ABL) and linear (FA) fits to estimate the quantities in Fig.~\ref{Fig10_supp}. \textbf{b}: Examples of ill-processed vertical profiles at 3 different time steps (blue - 2/12/2013 - 2300 UTC, orange - 20/03/2017 - 0000 UTC, green - 14/07/2017 - 1400 UTC) at the South Namib station. \textbf{c}: Hourly distribution of ill-processed vertical profiles. \textbf{d}: Monthly distribution of ill-processed vertical profiles.}
  \label{Fig9_supp}
\end{figure}
